\section{Product Motivation}

\begin{frame}{From Prediction to Decision}
  \centering
  \begin{tabular}{p{0.43\textwidth}p{0.43\textwidth}}
    \toprule
    \textbf{Prediction} & \textbf{Decision} \\
    \midrule
    What will happen? & What should we do? \\
    Truck arrival time & Which truck to unload next \\
    Grain quality & Where to route the grain \\
    Train delay & Which loading schedule to use \\
    \bottomrule
  \end{tabular}

  \vspace{0.65cm}
  {\large Machine learning describes the operation. Decision-making changes it.}

  \vspace{0.35cm}
  \normalsize
  Reinforcement learning is a framework for learning such decisions.
  \note{Start from familiar supervised learning. Prediction is useful for describing the operation, but the topic of the talk is choosing interventions: which truck, route, silo, or schedule. Keep RL as a light framing line only.}
\end{frame}

\begin{frame}{Why Decisions Are Different}
  \centering
  \begin{tikzpicture}[node distance=0.65cm,>=Latex,scale=0.92,transform shape]
    \node[draw,rounded corners,fill=Panel,minimum width=2.6cm,minimum height=0.85cm] (current) {current situation};
    \node[draw,rounded corners,fill=Panel,right=of current,minimum width=2.0cm,minimum height=0.85cm] (decision) {decision};
    \node[draw,rounded corners,fill=Panel,right=of decision,minimum width=2.7cm,minimum height=0.85cm] (changed) {changed operation};
    \node[draw,rounded corners,fill=Panel,right=of changed,minimum width=2.4cm,minimum height=0.85cm] (next) {next situation};
    \draw[->,thick] (current) -- (decision);
    \draw[->,thick] (decision) -- (changed);
    \draw[->,thick] (changed) -- (next);
  \end{tikzpicture}

  \vspace{0.55cm}
  \begin{itemize}
    \item Supervised learning describes a static world.
    \item Decision-making asks a counterfactual question: \emph{what should we do?}
    \item The chosen action changes future queues, inventories, blending options, and schedules.
  \end{itemize}

  \vspace{0.4cm}
  {\Large Prediction assumes a mostly stationary data source. \\ Decisions change the state distribution the model will face next.}
  \note{Make the key conceptual difference vivid without notation. Supervised learning usually describes the observed world: what label goes with this input? Decision-making asks about counterfactual futures: what should we do, and what would happen if we did it? Once deployed, the decision rule changes future state distributions, so the data source is no longer passive or stationary.}
\end{frame}

\begin{frame}{How Have We Solved Decision Problems?}
  \centering
  \begin{tikzpicture}[node distance=0.55cm,>=Latex]
    \node[draw,rounded corners,fill=Panel,minimum width=3.25cm,minimum height=1.15cm,align=center] (rules) {\textbf{Rules and}\\expert systems};
    \node[draw,rounded corners,fill=Panel,right=0.45cm of rules,minimum width=3.25cm,minimum height=1.15cm,align=center] (control) {\textbf{Optimization}\\and control};
    \node[draw,rounded corners,fill=Panel,right=0.45cm of control,minimum width=3.25cm,minimum height=1.15cm,align=center] (rl) {\textbf{Reinforcement}\\learning};

    \node[below=0.18cm of rules,align=center,text width=3.2cm] {\small Encode operational knowledge directly.};
    \node[below=0.18cm of control,align=center,text width=3.2cm] {\small Model the system and compute good actions.};
    \node[below=0.18cm of rl,align=center,text width=3.2cm] {\small Improve decisions using experience and feedback.};
  \end{tikzpicture}

  \vspace{1.55cm}
  These are not mutually exclusive: practical systems often combine expert knowledge, control, optimization, and learned policies.
  \note{Give a short historical orientation without turning it into a taxonomy. The point is that RL is one member of a larger decision-making family, not a mysterious replacement for control or optimization.}
\end{frame}

\begin{frame}{The Classical RL Idea}
  \centering
  \begin{tikzpicture}[node distance=0.55cm,>=Latex,scale=0.9,transform shape]
    \node[draw,rounded corners,fill=Panel,minimum width=1.8cm,minimum height=0.75cm] (observe) {observe};
    \node[draw,rounded corners,fill=Panel,right=of observe,minimum width=1.45cm,minimum height=0.75cm] (act) {act};
    \node[draw,rounded corners,fill=Panel,right=of act,minimum width=2.45cm,minimum height=0.75cm] (outcome) {see outcome};
    \node[draw,rounded corners,fill=Panel,right=of outcome,minimum width=1.75cm,minimum height=0.75cm] (improve) {improve};
    \node[draw,rounded corners,fill=Panel,right=of improve,minimum width=1.45cm,minimum height=0.75cm] (again) {act again};
    \draw[->,thick] (observe) -- (act);
    \draw[->,thick] (act) -- (outcome);
    \draw[->,thick] (outcome) -- (improve);
    \draw[->,thick] (improve) -- (again);
  \end{tikzpicture}

  \vspace{0.45cm}
  \Large
  Reinforcement learning learns a decision strategy through interaction.

  \vspace{0.25cm}
  \normalsize
  It can optimize long-term outcomes, account for how one decision affects later decisions, and learn strategies that are hard to specify manually.

  \vspace{0.3cm}
  But live trial and error is often expensive, slow, disruptive, or unsafe.
  \note{Introduce classical RL in plain language. The learner improves by interacting with the system. This is attractive because it can learn long-horizon behavior, but the practical limitation is live exploration. Avoid formal terms until the next slide.}
\end{frame}

\begin{frame}{A Modern Starting Point: Learn from Logged Decisions}
  \centering
  \begin{tabular}{p{0.43\textwidth}p{0.43\textwidth}}
    \toprule
    \textbf{Classical online RL} & \textbf{Offline approach} \\
    \midrule
    act & historical logs \\
    observe & train offline \\
    update & evaluate \\
    act again & cautious deployment \\
    \bottomrule
  \end{tabular}

  \vspace{0.55cm}
  Many real systems already contain years of operational decisions made by people, controllers, or earlier software.

  \vspace{0.3cm}
  \Large
  Offline RL asks whether we can learn a better decision strategy from logged interaction data, without new trial and error during training.
  \note{Frame offline RL as a modern data-driven formulation, not only a workaround for dangerous interaction. The key is starting from operational experience that already exists, then evaluating and rolling out cautiously.}
\end{frame}

\begin{frame}{Start Simple: Behavior Cloning}
  \centering
  \begin{tikzpicture}[node distance=0.55cm,>=Latex,scale=0.92,transform shape]
    \node[draw,rounded corners,fill=Panel,minimum width=3.15cm,minimum height=0.85cm,align=center] (data) {observed situation\\+ operator decision};
    \node[draw,rounded corners,fill=Panel,right=of data,minimum width=2.8cm,minimum height=0.85cm,align=center] (sl) {supervised\\learning};
    \node[draw,rounded corners,fill=Panel,right=of sl,minimum width=2.85cm,minimum height=0.85cm,align=center] (rule) {learned\\policy};
    \draw[->,thick] (data) -- (sl);
    \draw[->,thick] (sl) -- (rule);
  \end{tikzpicture}

  \vspace{0.35cm}
  \begin{itemize}
    \item Observe what an operator or existing controller did.
    \item Train a model to predict the same action in similar situations.
    \item Deploy the trained model as a learned ``policy''.
  \end{itemize}

  \vspace{0.25cm}
  Behavior cloning is fully offline and requires no simulator or reward model.

  \vspace{0.25cm}
  {\Large Can we do more than learn based on operator data? Can we prefer decisions that led to better long-term outcomes?}

  \vspace{0.2cm}
  To answer that, we first need a language for decision problems: states, actions, and policies.
  \note{This slide should make behavior cloning feel like the natural first step. It is just supervised learning on logged decisions. End by asking whether we can improve on the logs, then transition directly to the formal vocabulary slide.}
\end{frame}

\begin{frame}{States, Actions, and Policies}
  \centering
  \begin{tabular}{p{0.22\textwidth}p{0.65\textwidth}}
    \toprule
    \textbf{Object} & \textbf{Grain-terminal example} \\
    \midrule
    state $s_t$ & current queue, silo inventory, grain quality, train progress \\
    action $a_t$ & unload truck, choose pit, route conveyor, assign silo, fill car \\
    policy $\pi(a\mid s)$ & rule or model that chooses actions from states \\
    next state $s_{t+1}$ & the facility after the action has changed it \\
    \bottomrule
  \end{tabular}

  \vspace{0.55cm}
  A decision problem starts with states, actions, and a policy that maps situations to decisions.
  \note{Define the basic vocabulary before imitation learning uses it. State means the information available at decision time. Action means the decision we control. A policy is the rule or learned model that maps states to actions.}
\end{frame}

\begin{frame}{Agent-Environment Loop}
  \centering
  \begin{tikzpicture}[node distance=1.4cm,>=Latex]
    \node[draw,rounded corners,fill=Panel,minimum width=3cm,minimum height=1cm] (agent) {agent / policy $\pi$};
    \node[draw,rounded corners,fill=Panel,right=3.2cm of agent,minimum width=3cm,minimum height=1cm] (env) {environment};
    \draw[->,thick] (agent) -- node[above] {$a_t$} (env);
    \draw[->,thick] (env.south) -- ++(0,-1.0) -| node[pos=0.25,below] {$s_{t+1}, r_t$} (agent.south);
  \end{tikzpicture}

  \vspace{0.55cm}
  \begin{itemize}
    \item \textbf{Agent:} chooses the next action from the current state.
    \item \textbf{Environment:} the facility or simulation that returns the next state and reward.
    \item \textbf{Key contrast:} supervised learning predicts from fixed data; RL actions create the next data.
  \end{itemize}
  \note{This diagram clarifies the closed loop before imitation learning. The policy changes the data distribution by choosing actions. In this talk, the environment is either the physical grain facility or a simulator of that facility.}
\end{frame}
