\section{Idea 3: In-Sample Value Learning}

\begin{frame}{Problem: Can We Avoid Unsupported Maximization?}
  \Large
  Instead of correcting bad value estimates, can we avoid creating them?

  \vspace{0.4cm}
  \normalsize
  Many value-based methods rely on maximizing over actions:
  \[
    \max_a Q(s,a)
  \]

  \vspace{0.25cm}
  IQL asks a safer question:
  \[
    \text{What good actions are already supported by the data?}
  \]
  \note{This starts Chapter 3. The conceptual innovation is changing the question, not adding another loss function.}
\end{frame}

\begin{frame}{IQL: Replace the Max by a Supported Value}
  \large
  Many value-based methods ask for the best action through:
  \[
    \max_a Q(s,a).
  \]

  \vspace{0.2cm}
  IQL instead estimates:
  \[
    V(s)
    \approx
    \text{high expectile of } Q(s,a)
    \text{ for logged actions } a.
  \]

  \vspace{0.25cm}
  \begin{block}{Intuition}
    High but supported, instead of highest possible.
  \end{block}
  \note{This is the key mathematical contrast for IQL. Do not derive expectile regression; just explain it as a soft high value among actions in the data.}
\end{frame}

\begin{frame}{Central Idea: Learn Values Inside the Dataset}
  \centering
  \begin{tikzpicture}[>=Latex,x=0.95cm,y=0.9cm]
    \draw[->,thick] (0,0) -- (10,0) node[right] {actions in one state};
    \foreach \x/\h in {1.2/1.0,2.0/2.1,3.1/1.5,4.1/3.0,5.2/2.4,6.0/1.8,7.0/2.8} {
      \draw[fill=Panel] (\x-0.12,0) rectangle (\x+0.12,\h);
    }
    \draw[SignalGreen,ultra thick] (0.8,2.55) -- (7.4,2.55);
    \node[right] at (7.45,2.55) {high supported value};
    \draw[WarmGold,dashed,thick] (8.6,0) -- (8.6,3.7);
    \node[above] at (8.6,3.7) {unsupported action};
  \end{tikzpicture}

  \vspace{0.45cm}
  \Large
  IQL estimates high values among logged actions, not over all possible actions.

  \vspace{0.2cm}
  \normalsize
  In practice, this is implemented with expectile regression: a soft way to estimate a high supported value.
  \note{Mention expectiles once. Do not derive them. The audience should remember high-but-supported values.}
\end{frame}

\begin{frame}{Representative Method: IQL}
  \Large
  Implicit Q-Learning separates value estimation from action maximization.

  \vspace{0.35cm}
  \normalsize
  \begin{itemize}
    \item learn values from logged state-action pairs,
    \item avoid explicit maximization over unsupported actions,
    \item compute advantages using these in-sample values,
    \item train the policy with advantage-weighted behavior cloning.
  \end{itemize}

  \vspace{0.35cm}
  \begin{block}{What to remember}
    IQL = better value estimation + the same policy update as AWR.
  \end{block}
  \note{Close the loop back to AWR. IQL should feel like the answer to where the AWR weights come from.}
\end{frame}

\begin{frame}{IQL Connects Back to AWR}
  \large
  IQL changes how the advantage is computed:
  \[
    A(s,a)=Q(s,a)-V(s).
  \]

  \vspace{0.2cm}
  It does not change the policy update:
  \[
    w(s,a)=\exp\!\left(\beta A(s,a)\right),
    \qquad
    \min_\theta \sum_i w(s_i,a_i)\,
    \ell\!\left(\pi_\theta(s_i),a_i\right).
  \]

  \vspace{0.2cm}
  \begin{block}{Intuition}
    IQL gives better weights; the policy update is still weighted behavior cloning.
  \end{block}
  \note{This is one of the most important conceptual connections. IQL is not a completely different policy update; it improves the value estimates used by the AWR-style update.}
\end{frame}

\begin{frame}{Remaining Problem: What Should the Policy Look Like?}
  \Large
  The first three ideas focus on value estimation.

  \vspace{0.35cm}
  \normalsize
  But policy representation can also matter.

  \vspace{0.25cm}
  If several distinct actions are good, a simple policy may average them into a poor action.

  \vspace{0.35cm}
  \begin{block}{Optional next question}
    How should we represent policies when several actions are equally good?
  \end{block}
  \note{This sets up IDQL as complementary, not as a fourth big idea.}
\end{frame}
