\section{Offline RL}

\begin{frame}{Offline RL Setting}
  \Large
  Fixed dataset:
  \[
    \dataset = \{(s_t,a_t,r_t,s_{t+1})\}
  \]

  \vspace{0.2cm}
  Behavior policy:
  \[
    \pi_b(a \mid s)
  \]

  \vspace{0.2cm}
  Goal: learn $\pi(a \mid s)$ without online interaction during training.

  Logged data source: terminal operator logs. Action space: truck selection, pit assignment, conveyor route, silo, and loading schedule.
  \note{Define offline RL cleanly. The data was collected by a behavior policy pi-b, often not known exactly. Emphasize that this is policy improvement under support constraints, not supervised learning.}
\end{frame}

\begin{frame}{Support Mismatch}
  \centering
  \begin{tikzpicture}[>=Latex]
    \draw[->,thick] (0,0) -- (10,0) node[right] {action};
    \foreach \x/\lab in {2/$a_1$,3.2/$a_2$,4.3/$a_3$} {
      \fill[SignalGreen] (\x,0) circle (4pt);
      \node[below=0.18cm] at (\x,0) {\lab};
    }
    \node[draw,rounded corners,fill=Panel,above=0.8cm] at (3.15,0) {seen in $\dataset$};
    \fill[WarmGold] (8.1,0) circle (5pt);
    \node[below=0.18cm] at (8.1,0) {$a^\star$};
    \node[draw,rounded corners,fill=Panel,above=0.8cm] at (8.1,0) {chosen by learned policy};
  \end{tikzpicture}

  \vspace{0.6cm}
  The policy may choose an unsupported action where the value estimate is unreliable.
  \note{This is the core offline RL failure mode. The dataset covers actions near the behavior policy pi-b, but maximization may select an action far away.}
\end{frame}

\begin{frame}{Why Naive Q-Learning Fails Offline}
  \Large
  Dangerous backup:
  \[
    \max_a Q(s,a)
  \]

  \vspace{0.4cm}
  \normalsize
  The max can select actions that were rarely or never observed.

  \vspace{0.4cm}
  Then the learned policy may exploit errors in $Q$ outside the data distribution.
  \note{Explain overestimation: if the value function is wrong somewhere, the max operation will actively search for that mistake.}
\end{frame}

\begin{frame}{The Offline RL Thesis}
  \Large
  Offline RL is mostly about preventing the policy from exploiting value errors outside the data distribution.

  \vspace{0.5cm}
  \normalsize
  The rest of the talk builds a ladder:

  \[
    \text{BC} \rightarrow \text{AWR} \rightarrow \text{IQL} \rightarrow \text{IDQL}
  \]
  \note{Use this as the transition into algorithms. Each method fixes a limitation of the previous one.}
\end{frame}
