\section{Product Checklist}

\begin{frame}{What Matters in Real Offline RL Systems?}
  \small
  \begin{enumerate}
    \item \textbf{State quality:} is the decision approximately Markovian?
    \item \textbf{Action support:} does $\dataset$ cover actions we want $\pi$ to take?
    \item \textbf{Reward design:} does reward reflect throughput, quality, delay, rehandling, and train-loading penalties?
    \item \textbf{Counterfactual risk:} are we evaluating rare or unseen actions?
    \item \textbf{Evaluation:} offline evaluation, simulation, shadow deployment, or cautious rollout?
    \item \textbf{Policy constraints:} how far may $\pi$ move from $\pi_b$?
  \end{enumerate}

  \vspace{0.2cm}
  Safe deployment strategy: shadow recommendations, operator approval, and constrained rollout.
  \note{This gives the business and product audience a practical checklist. It also makes clear that offline RL is a systems problem, not just an algorithm.}
\end{frame}

\begin{frame}{Product Architecture Sketch}
  \centering
  \begin{tikzpicture}[node distance=0.85cm,>=Latex]
    \node[draw,rounded corners,fill=Panel,minimum width=2.8cm] (logs) {logged data};
    \node[draw,rounded corners,fill=Panel,right=of logs,minimum width=2.8cm] (train) {offline training};
    \node[draw,rounded corners,fill=Panel,right=of train,minimum width=2.8cm] (eval) {offline eval};
    \node[draw,rounded corners,fill=Panel,below=of eval,minimum width=2.8cm] (shadow) {shadow deploy};
    \node[draw,rounded corners,fill=Panel,left=of shadow,minimum width=2.8cm] (rollout) {safe rollout};
    \node[draw,rounded corners,fill=Panel,left=of rollout,minimum width=2.8cm] (monitor) {monitor KPIs};
    \draw[->,thick] (logs) -- (train);
    \draw[->,thick] (train) -- (eval);
    \draw[->,thick] (eval) -- (shadow);
    \draw[->,thick] (shadow) -- (rollout);
    \draw[->,thick] (rollout) -- (monitor);
    \draw[->,thick,bend right=35] (monitor) to (logs);
  \end{tikzpicture}

  \note{Use this as the product systems view. Offline RL does not remove the need for evaluation and careful rollout.}
\end{frame}

\begin{frame}{Closing Thesis}
  \Large
  RL is data-driven optimal control.

  \vspace{0.5cm}
  Offline RL is controlled policy improvement from logged data.

  \vspace{0.5cm}
  \normalsize
  The ladder:
  \[
    \text{control} \rightarrow \text{RL} \rightarrow \text{offline RL}
    \rightarrow \text{BC} \rightarrow \text{AWR} \rightarrow \text{IQL} \rightarrow \text{IDQL}
  \]
  \note{Close by returning to the core thesis. The audience should remember the ladder even if they forget the expectile objective.}
\end{frame}
